Sparsity Weak Lensing $3$-D Density Map Reconstruction: Reducing the line-of-sight smearing.

### 2020-10-23 doc/paper_ms_method_HSCY1/ms.tex
paper update

\begin{document}
\title{Sparsity Weak Lensing $3$-D Density Map Reconstruction}
%\author{Xiangchong Li}
%\affiliation{Department of Physics, University of Tokyo, Tokyo 113-0033, Japan}
%\affiliation{Kavli Institute for the Physics and Mathematics of the Universe (WPI),\\
%University of Tokyo, Kashiwa 277-8583, Japan}
%\author{Masamune Oguri}
%\affiliation{Department of Physics, University of Tokyo, Tokyo 113-0033, Japan}
%\affiliation{Kavli Institute for the Physics and Mathematics of the Universe (WPI),\\
%University of Tokyo, Kashiwa 277-8583, Japan}
%\affiliation{Research Center for the Early Universe, University of Tokyo, Tokyo 113-0033, Japan}
%\author{Wentao Luo}
%\affiliation{Kavli Institute for the Physics and Mathematics of the Universe (WPI),\\
%University of Tokyo, Kashiwa 277-8583, Japan}
%\author{HSC Collaboration}
%\noaffiliation
%\email{xiangchong.li@ipmu.jp}

\begin{abstract}
A new method is developed to reconstruct $3$-D density contrast maps from photometric weak lensing shear measurements.
The $3$-D density contrast maps is modeled as a summation of basis atoms which are designed to represent the profiles of 
multi-scale NFW halos. The NFW atoms have $2$-D multi-scale NFW surface density profiles on the transverse plane and 
$1$-D Dirac delta functions in the line-of-sight direction. With the sparsity prior assumption, the density field 
is reconstructed using an oracle algorithm: adaptive lasso. Our method is tested with realistic simulations using HSC-like 
shape noise and photo-$z$ uncertainties.
{\redColor{Add descriptions on the outcomes of the test on the simulations.}}
\end{abstract}

@@

\begin{figure*}[!ht]
    \includegraphics[width=0.98\textwidth]{nfwlet-atom-2D.pdf}
    \caption{The smoothed basis atoms with different scale radius ($r_s$). The first row shows the smoothed basis atoms in 
            Fourier space and the second row shows the smoothed basis atoms in Real space. The smoothing kernel is Gaussian
            with scale of $1.5$ pixels.} \label{fig-atoms2D}
\end{figure*}

\begin{figure}
 \includegraphics[width=0.45\textwidth]{nfwlet-atom-1D.pdf}
 \caption{The $1$-D slices for smoothed basis atoms at $x=0$. The corresponding $2$-D profiles are shown in Figure 
            \ref{fig-atoms2D}.}
\end{figure}

@@

Adopting the thin lens approximation, the expected shear measurements ($\gamma$) on distant galaxies are related to the 
foreground density contrast field ($\delta$) through a linear transformation
\begin{equation} \label{eq-intro-delta2shear}
 \gamma=\mathbf{T} \delta,
\end{equation}
where $\mathbf{T}$ is used to denote the linear transformation operator which includes not only physical lensing effect
but also observational systematic effects from the observation (e.g. pixelization and smoothing of shear field in
transverse plane, photo-$z$ uncertainties).

@@

Reconstructions of density map from shear measurements also receive considerable interest. $2$-D density map reconstruction 
which recover an integration of projected mass along the line-of-sight has been well studied within the community
\citep{massMap-KS1993,WL-massMap-Glimpse2D-Lanusse2016}
and applied to large scale surveys \citep{HSC1-massMaps,massMapDES-Chang2018,DES-SV-massMap-sparsity}. However, the 
reconstruction of $3$-D mass map is still a challenging task.

In order to fully reconstruct the $3$-D mass density distribution ($\delta$) from the tomographic shear observations ($\gamma$),
the density contrast field is modeled as a summation of basis atoms in a model dictionary
\begin{equation} \label{eq-intro-dict}
 \delta= \mathbf{\Phi} x,
\end{equation}
where $\mathbf{\Phi}$ is the transformation operator from the parameters in the dictionary space to the density contrast 
and $x$ denotes the parameters. \citet{LSS-massMap-Wiener-Simon2009} reconstruct the density field
in Fourier space, which is equivalent to model the mass field with sinusoidial functions. On the other hand,
\citet{LSS-massMap-Glimpse3D-Leonard2014} models the mass field with Starlets \citep{Starlet-Starck2015}.

The model parameters are estimated through solving an optimization problem and the estimator is defined as
\begin{equation}
\hat{x}=\argmin_{x} \left\{ \frac{1}{2}\norm{\Sigma^{-\frac{1}{2}}(\gamma- \mathbf{T\Phi} x)}_2^2+ \lambda C(x) \right\},
\end{equation}
where $\norm{\Sigma^{-\frac{1}{2}}(\gamma - \mathbf{T\Phi} x)}_2^2$ is the chi-square term\footnote{weighted by the
inverse of the diagonal covariance matrix of error on the shear measurements ($\Sigma$).} measuring
the residuals between the prediction and the data, while $C(x)$ is the regularization term measuring the deviation of
the estimation of the parameter ($x$) from the prior assumptions. Such estimation prefer the parameters that are able
to describe the observations and also align with the prior assumptions. The regularization parameter $\lambda$ adjusts 
the relative weight between the observations and prior assumptions in the optimization process.

\citet{LSS-massMap-Wiener-Simon2009} propose to use the $l^2$ ridge regulation ($C=\norm{x}^2_2$) to find a sparse 
solution in Fourier space. \citet{HSC1-massMaps} apply the method of \citet{LSS-massMap-Wiener-Simon2009} to the first 
year data of the Hyper Suprime-Cam Survey \citep{HSC1-data}. 
{\redColor{Why we use $l^1$ in this paper, what is the difference between $l^1$ and $l^2$.}}

\citet{LSS-massMap-Glimpse3D-Leonard2014} propose to use a derivative version of $l^1$ lasso regulation ($C=\norm{x}^1_1$) 
to find a sparse solution in the Starlets dictionary space \citep{Starlet-Starck2015}. However, the Starlets functions are 
not designed to model the profile of clumpy mass in the universe. Moreover, \citet{LSS-massMap-Glimpse3D-Leonard2014} 
apply a greedy coordinate descent algorithm, which selects the steepest coordinate in each iteration, find the minima of 
a non-convex loss function. However, the stability of the non-convex optimization has not been fully discussed.

$N$-body simulations have shown that the dark matter is distributed in halos connected by filaments, and the density 
profile of a single halo follows the NFW function \citep{halo-NFW1997ApJ}.
We construct a dictionary which is composed of multi-scale NFW atoms. The multi-scale NFW atoms are used to model the 
profiles of halos. The atoms follow multi-scale surface density profiles of NFW functions \citep{haloModel-TJ2003-3pt}
on the transverse plane.
The depth of halo is neglected \citep{LSS-massMap-Glimpse3D-Leonard2014} since the resolution scale of the reconstruction 
in the line-of-sight direction is much larger than the halo scale. Therefore, the Dirac delta function is used to model the
atoms' profile in the line-of-sight direction. 
With the prior assumption that the density contrast field is sparse in the dictionary space, the adaptive lasso regularization
\citep{AdaLASSO-Zou2006} is used to reconstruct a high signal-to-noise-ratio (SNR) density contrast field.

Comparing with \citet{LSS-massMap-Glimpse3D-Leonard2014}, our dictionary is built up to describe the clumpy mass 
in the universe which has a clear physical motivation. Furthermore, the adaptive lasso algorithm is full convex and 
can be optimized with the FISTA algorithm \citep{FISTA-Beck2009} without relying on any greedy coordinate descent approaches. 
The stability of this convex optimization has been well studied.

This paper is organized as follows.
Section \ref{sec:Method} proposes the new method for $3$-D density map reconstruction.
Section \ref{sec:Sim} introduces the realistic simulations we use to test the new method.
Section \ref{sec:Res} presents the results of our method on the simulations.
Section \ref{sec:Sum} summarizes and discusses the future development of the method.

@@

@@

Finally, we find the the sparse solution in section \ref{subsec:method-reversion} using the adaptive lasso algorithm 
\citep{AdaLASSO-Zou2006} optimized with the FISTA alogrithm \citep{FISTA-Beck2009}.

@@

The lensing convergence map at the comoving distance $\chi_s$ caused by the foreground inhomogeneous
density distribution at the comoving distance $\chi_l$ ($\chi_l< \chi_s$) along the line-of-sight is
\citep{LSS-massMap-Glimpse3D-Leonard2014}
\begin{equation}
\kappa(\vec{\theta},\chi_s)=\frac{3H_0^2\Omega_M}{2 c^2} \int_0^{\chi_s} d\chi_l \frac{\chi_l \chi_{sl}}{\chi_s}
\frac{\delta(\vec{\theta},\chi_l)}{a(\chi_l)},
\end{equation}
where $\delta=\rho(\vec{\theta},\chi_l)/\bar{\rho}-1$ is the density contrast
at the position of lens, $H_0$ is the Hubble parameter, $\Omega_M$ is the matter density parameter, $c$ is the speed
of light, and $a(\chi_l)$ is the scale parameter at the lens position.

Substitute comoving distance ($\chi$) with redshift ($z$)
\begin{equation}\label{eq-delta2kappa}
\kappa(\vec{\theta},z_s)=\int_0^{z_s} dz_l K(z_l,z_s)\delta(\vec{\theta},z_l).
\end{equation}
where $K(z_l,z_s)$ is the lensing kernel defined as
\begin{equation}
K(z_l,z_s) =
\begin{cases}
\frac{3H_0\Omega_M}{2 c} \frac{\chi_l \chi_{sl} (1+z_l)}{\chi_{s} E\left(z_l\right)} & (z_s>z_l),\\
0&(z_s \leq z_l).
\end{cases}
\end{equation}

As shown in \citet{massMap-KS1993}, the shear distortion is related to the kappa field at the redshift plane as
\begin{equation}\label{eq-kappa2gamma}
\gamma_L(\vec{\theta},z_s) = \int  d^2 \theta' D(\vec{\theta}-\vec{\theta'}) \kappa(\vec{\theta'},z_s),
\end{equation}
where
\begin{equation}
D(\vec{\theta})=-\frac{1}{\pi}(\theta_1-i\theta_2)^{-2}.
\end{equation}
Here we denote the physical shear distortions as $\gamma_L$ and we note that the observed shear measurements are the
physical shear distortions influenced by systematics which will be discussed in the following subsections.

Combining equation (\ref{eq-delta2kappa}) with equation (\ref{eq-kappa2gamma}), the physical shear signal is derived as
\begin{equation}\label{eq-delta2gammat}
\gamma_L(\vec{\theta},z_s) = \int_0^{z_s} dz_l K(z_l,z_s) \int d^2 \theta' \vec{D}(\vec{\theta}-\vec{\theta'}) \delta(\vec{\theta'},z_l).
\end{equation}

@@

The density contrast field is modeled as a summation of basis atoms in the dictionary space $\{\phi_0,\phi_1,...\phi_{N}\}$.
\begin{equation}\label{eq-x2delta}
\delta(\vec{r}) = \sum_{s=0}^{N} \int d^3 r' \phi_s(\vec{r}-\vec{r'}) x_s(\vec{r'}),
\end{equation}
where $\phi_s(\vec{r})$ is the basis vector of the dictionary space, and $x_s$ is the corresponding projection
coefficients of the density contrast field onto the basis vectors.

These basis vectors are composed of multi-scale NFW atoms $\{\phi_1,..,\phi_{N}\}$ and point
mass atoms $\{\phi_0\}$. $\phi_0$ is a $3$-D Dirac delta function
\begin{equation}
\phi_0(\vec{r})= \delta_D(\theta_1) \delta_D(\theta_2) \delta_D(z).
\end{equation}
on the transverse plane, the NFW atoms ($\{\phi_1,...,\phi_N\}$) follow surface density profiles of NFW halos
\citep{haloModel-TJ2003-3pt} with scale radius $\theta_\alpha$ and truncation radius $c \theta_\alpha$ ,
where $c$ is known as concentration of NFW halo.
As the scale of halo is much less than the reachable redshift resolution, we neglect the depth of halo on the
line-of-sight direction and set the profiles of NFW atoms in the line-of-sight direction to $1$-D Dirac delta functions
\citep{LSS-massMap-Glimpse3D-Leonard2014}. The NFW atoms are expressed as
\begin{equation}
\begin{split}
\phi_\alpha(\vec{r}) =&\frac{f }{2 \pi \theta_\alpha^2 } F(|\vec{\theta}|/\theta_\alpha) \delta_D(z),\\
&  (s=1..N)
\end{split}
\end{equation}
where
\begin{equation}
F(x)=
\begin{cases}
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} + \frac{\arccosh \left(\frac{x^2+c}{x(1+c)}\right)}{(1-x^2)^{3/2}}  & (x<1),\\
\frac{\sqrt{c^2-1}}{3(1+c)} (1+\frac{1}{c+1}) & (x=1),\\
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} + \frac{\arccos\left(\frac{x^2+c}{x(1+c)}\right)}{(x^2-1)^{3/2}} & (1<x\leq c),\\
0& (x>c).
\end{cases}
\end{equation}
$f=1/[\ln (1+c)-c/(1+c)]$. In this work, we fix $c=9$ for NFW atoms in different scale frames.

@@

Our prior assumption is that the mass density field is sparse in the dictionary space and we will conduct
adaptive lasso \citep{AdaLASSO-Zou2006} to find a sparse solution of the mass reconstruction problem in the
dictionary space.

We note that the shear measurements are deviated from the physical shear prediction due to the systematics
from observations. The influence of systematics is carefully studied and added to the forward modeling in the
following subsections.

@@

@@

\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth]{photo-z_binning.pdf}
 \caption{The source galaxies are binned into $10$ redshift bins according to their mlz best photo-$z$ estimation.
        The blue histogram is the number distribution of the best photo-$z$ estimation. The vertical dashed lines
        are the bounds of bins. The galaxies are evenly distributed in each bins.}
\end{figure}

Since the photometric redshifts of source galaxies in the current large scale survey are estimated with
a limited number of photometric bands, the estimated redshifts of galaxies suffer from large uncertainties.
Such uncertainties smear the structure along the line-of-sight direction since a galaxy with a best fit photo-$z$
estimation of $z_s$ has a possibility of being actually located at redshift $z$. Such probability is denoted as $P(z|z_s)$
and the expected shear distortion on this galaxy is
\begin{equation}\label{eq-delta2gamma-poz}
\gamma_L(\vec{\theta},z_s) \rightarrow \int dz_s P(z|z_s) \gamma_L(\vec{\theta},z_s).
\end{equation}
With the definition of photo-$z$ smearing operator
\begin{equation}
\mathbf{P} = \int dz_s P(z|z_s),
\end{equation}
the photo-$z$ uncertainties changes the shear as $\gamma_L \rightarrow \mathbf{Q} \gamma_L$.

@@

The observed galaxies have random irregular (unequally-spaced) spacial distribution. To benefit the computational
speed, we smooth the shear measurements from galaxy shapes and pixelise the smoothed measurements onto regular grids.

The smoothing is conducted by convolving the measurements with a smoothing kernel
\begin{equation}
\hat{\gamma}  = \frac{\sum_i  W(\vec{\theta}-\vec{{\theta}}_i,z-z_i) e_i}{\sum_i R_i W(\vec{\theta}-\vec{{\theta}}_i,z-z_i) },
\end{equation}
where $W(\vec{\theta},z)$ is a $3$-D smoothing kernel. $e_i$, $R_i$, $z_i$ and $\theta_i$ are the ellipticity,
response, reshift, and transverse position of the `$i$-th' galaxy in the galaxy catalog.

>>>>

\begin{document}
\title{Sparsity Weak Lensing $3$-D Density Map Reconstruction: 
Reducing the line-of-sight smearing.}
%\author{Xiangchong Li}
%\affiliation{Department of Physics, University of Tokyo, Tokyo 113-0033, Japan}
%\affiliation{Kavli Institute for the Physics and Mathematics of the Universe (WPI),\\
%University of Tokyo, Kashiwa 277-8583, Japan}
%\author{Masamune Oguri}
%\affiliation{Department of Physics, University of Tokyo, Tokyo 113-0033, Japan}
%\affiliation{Kavli Institute for the Physics and Mathematics of the Universe (WPI),\\
%University of Tokyo, Kashiwa 277-8583, Japan}
%\affiliation{Research Center for the Early Universe, University of Tokyo, Tokyo 113-0033, Japan}
%\author{Wentao Luo}
%\affiliation{Kavli Institute for the Physics and Mathematics of the Universe (WPI),\\
%University of Tokyo, Kashiwa 277-8583, Japan}
%\author{HSC Collaboration}
%\noaffiliation
%\email{xiangchong.li@ipmu.jp}

\begin{abstract}
A new method is developed to reconstruct $3$-D density contrast maps from photometric weak lensing shear measurements.
The $3$-D density contrast maps is modeled as a summation of NFW basis atoms which have $2$-D multi-scale NFW surface 
density profiles on the transverse plane and $1$-D Dirac delta functions in the line-of-sight direction. With the prior 
assumption that the density fields can be sparsely represented by the NFW basis atoms, it is reconstructed using an 
oracle algorithm: adaptive lasso. Our method is tested with realistic simulations using the HSC-like shape noise and 
photo-$z$ uncertainty.
{\redColor{Add descriptions on the outcomes of the test on the simulations.}}
\end{abstract}

@@

\begin{figure*}[!t]
    \includegraphics[width=1.\textwidth]{nfwlet-atom-2D.pdf}
    \caption{The smoothed basis atoms. The first row shows the smoothed basis atoms in Fourier space 
            and the second row shows the smoothed basis atoms in Real space. 
            The first column is point mass atom and the other columns are multi-scale NFW atoms.
            The smoothing kernel is Gaussian with scale of $1.5$ pixels.} \label{fig-atoms2D}
\end{figure*}

\begin{figure}
 \includegraphics[width=0.5\textwidth]{nfwlet-atom-1D.pdf}
 \caption{The $1$-D slices for smoothed basis atoms at $x=0$. The corresponding $2$-D profiles are shown in Figure 
        \ref{fig-atoms2D}.} \label{fig-atoms1D}
\end{figure}

@@

The expected shear measurements ($\gamma$) on distant galaxies are related to the 
foreground density contrast field ($\delta$) through a linear transformation
\begin{equation} \label{eq-intro-delta2shear}
 \gamma=\mathbf{T} \delta,
\end{equation}
where $\mathbf{T}$ is used to denote the linear transformation operator which includes not only physical lensing effect
but also systematic effects from observations (e.g. pixelization and smoothing of shear field in
transverse plane, photo-$z$ uncertainty).

@@

Reconstructions of density map from shear measurements also receive considerable interest. $2$-D density map reconstruction 
which recover an integration of projected mass along the line-of-sight has been well studied within the community
\citep{massMap-KS1993,WL-massMap-Glimpse2D-Lanusse2016,sparseBaysianMassMap-Price2020}
and applied to large scale surveys \citep{HSC1-massMaps,massMapDES-Chang2018,DES-SV-massMap-sparsity}. However, the 
reconstruction of $3$-D mass map is still a challenging task.

In order to fully reconstruct the $3$-D mass density distribution ($\delta$) from the photometric shear observations ($\gamma$),
the density contrast field is modeled as a summation of basis atoms in a model dictionary
\begin{equation} \label{eq-intro-dict}
 \delta= \mathbf{\Phi} x,
\end{equation}
where $\mathbf{\Phi}$ is the transformation operator from the parameters in the dictionary space to the density contrast 
and $x$ denotes the parameters. \citet{LSS-massMap-Wiener-Simon2009} reconstruct the density field
in Fourier space, which is equivalent to model the mass field with sinusoidial functions. On the other hand,
\citet{LSS-massMap-Glimpse3D-Leonard2014} models the mass field with Starlets \citep{Starlet-Starck2015}.

The projection coefficients are estimated through optimizing a loss function with regularization. The estimator is
generally defined as
\begin{equation}
\hat{x}=\argmin_{x} \left\{ \frac{1}{2}\norm{\Sigma^{-\frac{1}{2}}(\gamma- \mathbf{T\Phi} x)}_2^2+ \lambda C(x) \right\},
\end{equation}
where $\norm{\Sigma^{-\frac{1}{2}}(\gamma - \mathbf{T\Phi} x)}_2^2$ is the chi-square term\footnote{weighted by the
inverse of the diagonal covariance matrix of error on the shear measurements ($\Sigma$).} measuring
the residuals between the prediction and the data, while $C(x)$ is the regularization term measuring the deviation of
the estimation of the parameter ($x$) from the prior assumptions. Such estimation prefers the parameters that are able
to describe the observations and also align with the prior assumptions. The regularization parameter $\lambda$ adjusts 
the relative weight between the observations and prior assumptions in the optimization process.

\citet{LSS-massMap-Wiener-Simon2009} propose to use the Wiener filter, which is also known as $l^2$ ridge regulation 
($C=\norm{x}^2_2$), to find a solution in Fourier space. \citet{HSC1-massMaps} apply the method of 
\citet{LSS-massMap-Wiener-Simon2009} to the first year data of the Hyper Suprime-Cam Survey \citep{HSC1-data}. 
However, the density maps reconstructed by this method suffer from a smearing along the line-of-sight direction with 
a standard deviation of $\sigma_z=0.2 \sim 0.3$.

\citet{LSS-massMap-Glimpse3D-Leonard2014} propose to use a derivative version of $l^1$ lasso regulation ($C=\norm{x}^1_1$) 
to find a sparse solution in the Starlets dictionary space \citep{Starlet-Starck2015}. \citet{LSS-massMap-Glimpse3D-Leonard2014}
apply a greedy coordinate descent algorithm, which selects the steepest coordinate in each iteration, to find the minima 
of a non-convex loss function penalized with the firm thresholding function. \citet{LSS-massMap-Glimpse3D-Leonard2014} 
significantly reduce the smearing along the line-of-sight, however, the stability of the non-convex optimization and
greedy coordinate descent algorithm has not been fully justified. Moreover, the Starlets functions are not designed to 
model the profile of clumpy mass in the universe.

$N$-body simulations have shown that the dark matter is distributed in halos connected by filaments, and the density 
profile of a single halo follows the NFW function \citep{halo-NFW1997ApJ}.
We construct a dictionary with multi-scale NFW atoms.  The atoms follow multi-scale surface density profiles of NFW 
functions \citep{haloModel-TJ2003-3pt} on the transverse plane. We follow \citet{LSS-massMap-Glimpse3D-Leonard2014} 
to neglect the depth of halos since the resolution scale of the reconstruction in the line-of-sight direction is much 
larger than the scales of halos. Therefore, we set the profile of the NFW atoms in the line-of-sight as the Dirac delta 
function. The multi-scale NFW atoms are used to model the profiles of halos which is assumed to be sparsely distributed 
in the universe. With the sparsity prior, the adaptive lasso regularization \citep{AdaLASSO-Zou2006} is used to reconstruct 
the density field. We find that, in contrast to the standard lasso estimator which smears the structure along the line-of-sight,
the addaptive lasso is able to significantly reduce the smearing effect.

Comparing with \citet{LSS-massMap-Glimpse3D-Leonard2014}, our dictionary is built up to describe the clumpy mass 
in the universe which has a clear physical motivation. Furthermore, the adaptive lasso algorithm is strictly convex and 
can be directly optimized with the FISTA algorithm \citep{FISTA-Beck2009} without relying on any greedy coordinate 
descent approaches.  The stability of this convex optimization has been well studied.

This paper is organized as follows.
Section \ref{sec:Method} proposes the new method for $3$-D density map reconstruction.
Section \ref{sec:Test} Tests our algorithm on HSC-like simulations.
Section \ref{sec:Sum} summarizes and discusses the future development of the method.

@@

\begin{figure}[!t]
 \centering
 \includegraphics[width=0.5\textwidth]{photo-z_binning.pdf}
 \caption{The source galaxies are binned into $10$ redshift bins according to their MLZ best photo-$z$ estimation.
        The blue histogram is the number distribution of the best photo-$z$ estimation. The vertical dashed lines
        are the bounds of bins. The galaxies are evenly distributed in each bins.} \label{fig-bestpz}
\end{figure}

@@

Finally, we find the sparse solution in section \ref{subsec:method-reconstruction} using the adaptive lasso algorithm 
\citep{AdaLASSO-Zou2006} optimized with the FISTA alogrithm \citep{FISTA-Beck2009}.

@@

\begin{figure}[!t]
 \centering
 \includegraphics[width=0.5\textwidth]{mlz-poz.pdf}
 \caption{The average PDF of MLZ photo-$z$ error for $10$ source redshift bins.}\label{fig-pdfpz}
\end{figure}

The lensing convergence map at the comoving distance $\chi_s$ caused by the foreground inhomogeneous
density distribution at the comoving distance $\chi_l$ ($\chi_l< \chi_s$) along the line-of-sight is
\begin{equation}
\kappa(\vec{\theta},\chi_s)=\frac{3H_0^2\Omega_M}{2 c^2} \int_0^{\chi_s} d\chi_l \frac{\chi_l \chi_{sl}}{\chi_s}
\frac{\delta(\vec{\theta},\chi_l)}{a(\chi_l)},
\end{equation}
where $\delta=\rho(\vec{\theta},\chi_l)/\bar{\rho}-1$ is the density contrast
at the position of lens, $H_0$ is the Hubble parameter, $\Omega_M$ is the matter density parameter, $c$ is the speed
of light, and $a(\chi_l)$ is the scale parameter at the lens position.

Substitute comoving distance ($\chi$) with redshift ($z$), we have
\begin{equation}\label{eq-delta2kappa}
\kappa(\vec{\theta},z_s)=\int_0^{z_s} dz_l K(z_l,z_s)\delta(\vec{\theta},z_l).
\end{equation}
where $K(z_l,z_s)$ is the lensing kernel defined as
\begin{equation}
K(z_l,z_s) =
\begin{cases}
\frac{3H_0\Omega_M}{2 c} \frac{\chi_l \chi_{sl} (1+z_l)}{\chi_{s} E\left(z_l\right)} & (z_s>z_l),\\
0&(z_s \leq z_l),
\end{cases}
\end{equation}
where $E(z)$ is the Hubble parameter as a function of redshift, in units of $H_0$.

\begin{figure*}[!t]
 \centering
 \includegraphics[width=1.\textwidth]{lensing_kernel.pdf}
 \caption{The left panel shows the lensing kernels for five different lens redshifts. The dashed lines are the 
        kernels for spectroscopic redshift which assumes that the redshifts of source galaxies are precisely 
        estimated. The solid lines are for photometric redshifts which accounts for the influence of photometric 
        redshift uncertainty. The other two panels show the correlation between lensing kernels of different 
        lens redshifts. The middle panel is for spectroscopic redshift and the right panel is for photometric 
        redshifts. The lensing kernels are normalized so that the diagonal elements of the correlation matrices
        equal one.}\label{fig-corlensKer}
\end{figure*}

As shown in \citet{massMap-KS1993}, the shear field is related to the kappa field at the same redshift plane
via
\begin{equation}\label{eq-kappa2gamma}
\gamma_L(\vec{\theta},z_s) = \int  d^2 \theta' D(\vec{\theta}-\vec{\theta'}) \kappa(\vec{\theta'},z_s),
\end{equation}
where
\begin{equation}
D(\vec{\theta})=-\frac{1}{\pi}(\theta_1-i\theta_2)^{-2}.
\end{equation}
Here we denote the physical shear distortion field as $\gamma_L$ and we note that the final observed shear measurements 
are also physical shear distortions influenced by systematic errors from observatins. The systematic error will be discussed 
in Section \ref{subsec:method-Systematics}.

Combining equation (\ref{eq-delta2kappa}) with equation (\ref{eq-kappa2gamma}), the expectation of lensing shear signal is 
\begin{equation}\label{eq-delta2gammat}
\gamma_L(\vec{\theta},z_s) = \int_0^{z_s} dz_l K(z_l,z_s) \int d^2 \theta' \vec{D}(\vec{\theta}-\vec{\theta'}) \delta(\vec{\theta'},z_l).
\end{equation}

@@

\begin{figure*}[!t]
\centering
\includegraphics[width=0.42\textwidth]{delta-3-7-pz-nn-lasso.png}
\includegraphics[width=0.42\textwidth]{delta-3-7-pz-nn-alasso.png}
\caption{The result of density map reconstruction with standard lasso (left) and adaptive lasso (right). The input density 
        map is from a NFW halo with mass $M_{200}=10^{15} ~h^{-1}M_{\odot}$ at redshift $0.35$. The lasso reconstruction 
        smear the density field along the line-of-sight direction. Noises on shear measurements are neglected in the 
        simulation.} \label{fig-lassoVsadaLasso}
\end{figure*}

The density contrast field is modeled as a summation of basis atoms in the dictionary 
\begin{equation}\label{eq-x2delta}
\delta(\vec{r}) = \sum_{s=1}^{N} \int d^3 r' \phi_s(\vec{r}-\vec{r'}) x_s(\vec{r'}),
\end{equation}
where $\phi_s(\vec{r})$ are the basis atoms of the dictionary. The basis atoms have `$N$' different scales and the atoms
in each scale frame are shifted by $\vec{r'}$ to form basis at different positions. $x_s(\vec{r'})$ is the projection 
coefficients of the density contrast field onto the basis atoms.

These basis atoms are constructed with multi-scale NFW atoms which are denoted as $\{\phi_1,...,\phi_N\}$.
On the transverse plane, the NFW atoms follow surface density profiles of NFW halos
\citep{haloModel-TJ2003-3pt} with scale radius $\theta_\alpha$ and truncation radius $c \theta_\alpha$ ,
where $c$ is the concentration of NFW halos.
As the scales of halos are much less than the reachable redshift resolution, we neglect the depth of halo on the
line-of-sight direction and set the profiles of NFW atoms in the line-of-sight direction to $1$-D Dirac delta 
functions as suggested by \citep{LSS-massMap-Glimpse3D-Leonard2014}. 
The multi-scale NFW atoms are defined as
\begin{equation}
\begin{split}
\phi_\alpha(\vec{r}) =&\frac{f }{2 \pi \theta_\alpha^2 } F(|\vec{\theta}|/\theta_\alpha) \delta_D(z),\\
&  (s=1..N)
\end{split}
\end{equation}
where
\begin{equation}
F(x)=
\begin{cases}
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} + \frac{\arccosh \left(\frac{x^2+c}{x(1+c)}\right)}{(1-x^2)^{3/2}}  & (x<1),\\
\frac{\sqrt{c^2-1}}{3(1+c)} (1+\frac{1}{c+1}) & (x=1),\\
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} + \frac{\arccos\left(\frac{x^2+c}{x(1+c)}\right)}{(x^2-1)^{3/2}} & (1<x\leq c),\\
0& (x>c).
\end{cases}
\end{equation}
$f=1/[\ln (1+c)-c/(1+c)]$. In this work, we fix $c=4$ for NFW atoms in different scale frames.

@@

In this paper dictionary constructed with point mass atoms is used to compare with the dictionary of multi-scale atoms.
The point mass atoms is a $3$-D Dirac function defined as follows
\begin{equation}
\phi_{\rm{PM}}(\vec{r})= \delta_D(\theta_1) \delta_D(\theta_2) \delta_D(z).
\end{equation}

The $2$-D profiles of the point mass atom and the multi-scale NFW atoms on the transverse plane are shown in Figure 
\ref{fig-atoms2D}. The $1$-D slices of the profiles are demonstrated in Fiugre \ref{fig-atoms1D}. Note that these profiles 
are smoothed with a Gaussian kernel and pixelized into evenly spaced grids. The smoothing operator is described in 
Section \ref{subsec:method-smoothing} and the pixelization operator is described in Section \ref{subsec:method-pixel}.

@@

The observed shear measurements are deviated from the physical shear prediction due to the systematic 
errors from observations. The influence of systematics is carefully studied and incorporated into the forward 
modeling in this section.

@@

The photometric redshifts of source galaxies in the current large scale survey are estimated with a limited number of 
board photometric bands (e.g. $9$ bands for KIDS$+$VIKING survey, $5$ bands for DES survey and HSC survey). As a result, 
the estimated redshifts of galaxies suffer from much larger uncertainties, comparing with redshifts estimated with 
spectroscopic observations. Such photo-$z$ uncertainty smears the lensing kernels statistically since a galaxy with a 
best fit photo-$z$ estimation of $z_s$ has possibilities of being actually located at different redshifts ($z$).
The probability function for the photo-$z$ uncertainty is denoted as $P(z|z_s)$ and the expected shear distortion on 
the galaxy is
\begin{equation}\label{eq-delta2gamma-poz}
\gamma_L(\vec{\theta},z_s) \rightarrow \int dz_s P(z|z_s) \gamma_L(\vec{\theta},z_s).
\end{equation}
With the definition of photo-$z$ smearing operator
\begin{equation}
\mathbf{P} = \int dz_s P(z|z_s),
\end{equation}
the photo-$z$ uncertainty changes the shear as $\gamma_L \rightarrow \mathbf{Q} \gamma_L$.

Figure \ref{fig-bestpz} shows the histogram of the best MLZ photo-$z$ estimation \cite{HSC1-photoz} for galaxies in tract 
9347 of HSC S16A data release \citep{HSC1-data}. These Galaxies are divided into ten source galaxy bins according to the 
photo-$z$ best estimation and the boundaries of the bins are shown as vertical dash lines in Figure \ref{fig-bestpz}. Figure \ref{fig-pdfpz} shows the average probability density functions (PDFs) in each redshift bin.

The left panel of Figure \ref{fig-corlensKer} shows the lensing kernels for lenses at five different redshifts as functions 
of source galaxy redshfit bins. The dashed lines are the lensing kernel for spectroscopic redshfits with neglectable redshift 
uncertainty. The solid lines are the lensing kernel for photometric redshifts with redshift uncertainties shown in Figure 
\ref{fig-pdfpz}. The middle and right panels show the correlation between lensing kernels for lenses at different redshifts. 
The middle panel is for spect-$z$ and the right panel is for photo-$z$.

The photo-$z$ uncertainties smear the shapes of lensing kernels. As demonstrated by the dashed lines in the left panel of 
Figure \ref{fig-corlensKer}, the lensing kernels converge to zero for source redshifts lower than the lens redshift, if the 
uncertainties on the source galaxy redshift estimations are neglectable. However, as demonstrated by the solid lines in the 
same panel, for the source redshifts with large photo-$z$ uncertainties, the lensing kernels do not converge to zero at 
redshifts lower than the lens redshift. This is because the galaxies with photo-$z$ estimations lower than the lens redshifts 
may be actually located at higher redshifts due to the photo-$z$ uncertainties.
Comparing the correlation matrices shown in the middle panel and the right panel of Figure \ref{fig-corlensKer}, we conclude 
that the smearings of lensing kernels due to photo-$z$ uncertainties increase the correlations between lensing kernels at 
different lens plane.

@@

\begin{figure}[!t]
 \centering
 \includegraphics[width=0.45\textwidth]{shapeMeasurementError-HSCY1.pdf}
 \caption{The Histogram of HSC-like shear measurement error (including errors from shape noise and photon noise) on the 
        first component of shear ($g_1$).}
 \label{fig:mass-redshift}
\end{figure}

The observed galaxies have random irregular (unequally-spaced) spacial distribution. In order boost the computational
speed, we smooth the shear measurements from galaxy shapes and pixelize the smoothed measurements onto regular grids.
After the pixelization, the fast Fourier transform (FFT) can be directly conducted on the transverse plane in each 
source redshift bin.

The smoothing is conducted by convolving the shear measurements with a smoothing kernel
\begin{equation}
\gamma_{\rm{sm}} (\vec{\theta})  = \frac{\sum_i  W(\vec{\theta}-\vec{{\theta}}_i,z-z_i) \gamma_i}{\sum_i W(\vec{\theta}-\vec{{\theta}}_i,z-z_i) },
\end{equation}
where $W(\vec{\theta},z)$ is a $3$-D smoothing kernel. $\gamma_i$, $z_i$ and $\theta_i$ are the shear,
photometric reshift, and transverse position of the `$i$-th' galaxy in the catalog.
